A Study of Common Bug Fix Patterns in Rust \thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}

### 2023-05-27 discussion.tex
add two todos

The seventh-most-popular pattern is the Rust-specific change of dropping mutability (i.e. making data immutable). Tools like ReImInfer~\cite{HuangMDE2012} for a Java extension, and techniques like the one proposed by Eyolfson~\cite{EyolfsonAbstractImmutability} for C++, aim to infer that data is immutable. In the Rust context, it is easier to make data immutable: the semantics for immutability are clear. Immutability is part of the core Rust language. mmutability escape hatches in C++ complicate issues, but in Rust, the escape hatch is the clearly-marked \texttt{unsafe} block. This pattern thus suggests that tools for inferring Rust immutability could be viable.

>>>>

The seventh-most-popular pattern is the Rust-specific change of dropping mutability (i.e. making data immutable). Tools like ReImInfer by~\cite{HuangMDE2012} (for a Java extension), and techniques like the one proposed in~\cite{EyolfsonAbstractImmutability} for C++, aim to infer that data is immutable. In the Rust context, it is easier to make data immutable: the semantics for immutability are clear. Immutability is part of the core Rust language. mmutability escape hatches in C++ complicate issues, but in Rust, the escape hatch is the clearly-marked \texttt{unsafe} block. This pattern thus suggests that tools for inferring Rust immutability could be viable.

### 2023-05-27 main.tex
final commit before submission
add two todos

\institute{Mohammad Robati Shirzad \at
              200 University Ave W, Waterloo, ON N2L 3G1 \\
              \email{mrobatis@uwaterloo.ca}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
}

@@

Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs. In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust. Ruxanne implements a novel embedding of Rust code into fixed-sized vectors. Using Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories. We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns. We describe each of these patterns and their implications to automated program repair.
\keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{68N01 \and 68N15} %% Computer Science : General and Programming Languages
\end{abstract}

>>>>

\institute{Mohammad Robati Shirzad --- \email{mrobatis@uwaterloo.ca} \\
           Patrick Lam --- \email{patrick.lam@uwaterloo.ca} \at
           University of Waterloo --- Address: 200 University Ave W, Waterloo, ON N2L 3G1 \\
%             \emph{Present address:} of F. Author  %  if needed
}

@@

Yet, to the best of our knowledge, no research has sought to identify the most pervasive bug fix patterns within Rust programs. In this project, we introduce Ruxanne, a tool for analyzing and extracting fix patterns in Rust. Ruxanne implements a novel embedding of Rust code into fixed-sized vectors. Using Ruxanne, we mined the top 18 most-starred Rust projects in Github to discover the most common bug fix patterns committed to their repositories. We analyzed 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 20 groups of cross-project bug fix patterns, which we categorize as (1) general patterns and (2) borrow-checker-related patterns. We describe each of these patterns and their implications to automated program repair.\todo{talk about the most popular pattern}
\keywords{Bug patterns \and Pattern mining \and Bug fix changes \and Rust}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{68N01 \and 68N15} %% Computer Science : General and Programming Languages
\end{abstract}
